An Imaging Search for Post-MS Companions of Sirius B

### 2021-11-04 paper/paper.tex
update paper

There is limited knowledge of planetary systems around evolved stars. The pathway for a \textit{first-generation} planet around a post-MS host is violent and uncertain. For a planet to survive the giant branch evolution, it first needs to escape engulfment as well as tidal shredding \citep{burleighImagingPlanetsNearby2002a,nordhausOrbitsLowmassCompanions2013}. Surviving planets are subject to adiabatic orbit expansion \citep{jeansCosmogonicProblemsAssociated1924}, which can potentially de-stabilize multi-object systems, stellar winds, and high irradiation which enriches the circumstellar environment with metals and dust, and strongly heats planets close to the giant star \citep{mustillForetellingsRagnarokWorldengulfing2012,verasPostmainsequencePlanetarySystem2016}. The expansion and tidal effects on the circumstellar environment create a ``forbidden'' region of exoplanet phase space for orbital separations closer than $\sim$\qty{5}{\au}.

@@

High-contrast imaging is a powerful technique for discovering and characterizing exoplanets. Directly imaging a circumstellar companion is difficult, though. The typical astrophysical flux ratios (contrasts) for a Sun-Jupiter analog in the near-infrared (NIR) are $\sim$$10^{\text{-}8}$, and for a Sun-Earth system are $\sim$$10^{\text{-}10}$ \citep{traubDirectImagingExoplanets2010}. In addition, the close angular separations of planets make it difficult to detect them over the diffraction pattern of their host and other noise sources. Thermal emission from exoplanets peaks, well into in the infrared, on the Rayleigh-Jeans limit of the stellar spectrum. This makes infrared observations preferred over optical or ultraviolet observations due to the reduced contrast.

WDs are interesting targets for direct imaging- they are \numrange{3}{4} orders of magnitude fainter than their MS progenitors which reduces the contrast necessary to detect a companion. The high effective temperatures of WD photospheres pushes the peak thermal emission even further into the optical and ultraviolet, reducing the intrinsic contrast between the star and the planet. In addition, the lack of spectral features of WDs \citep{schatzmanTheorieDebitEnergie1945} makes them imprecise targets for the radial velocity technique. The faintness of WDs proves challenging due to the high signal-to-noise ratio (S/N) required by modern adaptive optics (AO) instruments. This is exacerbated on ground-based telescopes which require extreme AO to counteract the effects of atmospheric seeing. This constrains potential companion searches to relatively bright targets ($m^R\lesssim$ 13). Implicitly, this also limits searches to nearby WDs ($d\sim$\qty{25}{\parsec}, \citealp{holberg25ParsecLocal2016}), which is beneficial for imaging because the limited inner working angle of the instrumentation will project to closer separations around closer stars.

@@

One of the peculiarities of the Sirius system is its large eccentricity of 0.59 \citep{bondSiriusSystemIts2017}. If we assume the orbital expansion due to Sirius B's evolution was adiabatic, we can calculate the initial semi-major axis of the binary
\begin{equation}
    a_i = a_f \frac{M_{B,f} + M_{A,f}}{M_{B,i} + M_{A,i}}
\end{equation}
where $a$ is the system semi-major axis, $M_A$ and $M_B$ are the respective stellar masses, and subscripts $i$ and $f$ correspond to the initial (MS) versus final (post-MS) states. The current semi-major axis of the binary is \qty{20}{\au}, and assuming negligible mass transfer between the two stars, the initial semi-major axis would be \qty{8.6\pm1.3}{\au}. If the orbit expansion was indeed adiabatic, the eccentricity would be the same before and after evolution. In this case, the periastron of Sirius A and B would be \qty{3.52\pm0.52}{\au}. \citet{verasPostmainsequencePlanetarySystem2016} tabulated the maximum stellar radius of intermediate mass stars during their giant evolution, from which we interpolate a maximum radius for Sirius B of \qty{5.104\pm0.075}{\au}. This means Sirius A certainly interacted with Sirius B and may have had a common envelope stage. Mass transfer and tidal circularization would be expected, however the present-day eccentricity provide contrary evidence.

@@

\section{Observations}\label{sec:obs}
\defcitealias{bonnet-bidaudADONISHighContrast2008a}{BB08}
\defcitealias{bondSiriusSystemIts2017}{B17}
\defcitealias{collaborationGaiaEarlyData2021a}{G21a}
\defcitealias{cummingsWhiteDwarfInitialFinal2018}{C18}
\begin{deluxetable}{cccc}
    \tablecaption{Parameters of the Sirius system adopted in this study.\label{tbl:system}}
    \tablehead{
        \colhead{parameter} & \colhead{value} & \colhead{unit} & \colhead{ref.}
    }
    \startdata
    $t_\mathrm{sys}$ & 225 & \unit{\mega\year} & \citetalias{bondSiriusSystemIts2017,cummingsWhiteDwarfInitialFinal2018} \\
    $\pi$ & \num{374.49\pm0.23} & mas & \citetalias{collaborationGaiaEarlyData2021a} \\
    $d$ & \num{2.6702\pm.0016} & pc & \citetalias{collaborationGaiaEarlyData2021a} \\
    $a$ & \num{20.016\pm0.014} & \unit{\au} & \citetalias{bondSiriusSystemIts2017} \\
    \cutinhead{Sirius A}
    $M_\star$ & \num{2.063\pm0.023} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017} \\
    \cutinhead{Sirius B}
    $M_\star$ & \num{1.018\pm0.011} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017} \\
    $M_\mathrm{MS}$ & \num{5.1\pm1.1} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017,cummingsWhiteDwarfInitialFinal2018} \\
    $t_\mathrm{WD}$ & 126 & \unit{\mega\year} & \citetalias{bondSiriusSystemIts2017}\\
    $m^{\mathrm{L}^\prime}$ & \num{9.1\pm0.2} & & \citetalias{bonnet-bidaudADONISHighContrast2008a} \\
    $M^{\mathrm{L}^\prime}$ & \num{11.97\pm0.2} & & \citetalias{bonnet-bidaudADONISHighContrast2008a,collaborationGaiaEarlyData2021a} \\
    \enddata
\end{deluxetable}

>>>>

There is limited knowledge of planetary systems around evolved stars. The pathway for a \textit{first-generation} planet around a post-MS host is violent and uncertain. For a planet to survive the giant branch evolution, it first needs to escape engulfment as well as tidal shredding \citep{burleighImagingPlanetsNearby2002a,nordhausOrbitsLowmassCompanions2013}. Surviving planets are subject to adiabatic orbit expansion, which can potentially de-stabilize multi-object systems, stellar winds, and high irradiation which enriches the circumstellar environment with metals and dust, and strongly heats planets close to the giant star \citep{mustillForetellingsRagnarokWorldengulfing2012,verasPostmainsequencePlanetarySystem2016}. The expansion and tidal effects on the circumstellar environment create a ``forbidden'' region of exoplanet phase space for orbital separations closer than $\sim$\qty{5}{\au}.

@@

High-contrast imaging is a powerful technique for discovering and characterizing exoplanets. Directly imaging a circumstellar companion is difficult, though. The typical astrophysical flux ratios (contrasts) for a Sun-Jupiter analog in the near-infrared (NIR) are $\sim$$10^{\text{-}8}$, and for a Sun-Earth system are $\sim$$10^{\text{-}10}$ \citep{traubDirectImagingExoplanets2010}. In addition, the close angular separations of planets make it difficult to detect them over the diffraction pattern of their host and other noise sources. Thermal emission from exoplanets peaks well into the infrared, on the Rayleigh-Jeans limit of the stellar spectrum. This makes infrared observations preferred over optical or ultraviolet observations due to the reduced contrast.

WDs are interesting targets for direct imaging- they are \numrange{3}{4} orders of magnitude fainter than their MS progenitors which reduces the contrast necessary to detect a companion. The high effective temperatures of WD photospheres pushes the peak thermal emission even further into the optical and ultraviolet, reducing the intrinsic contrast between the star and the planet. In addition, the lack of spectral features of WDs \citep{schatzmanTheorieDebitEnergie1945} makes them imprecise targets for the radial velocity technique. The small stellar radii of WDs lowers the transit probability significantly, although detections of substellar companions around WDs have been made \citep[e.g.,][]{vanderburgGiantPlanetCandidate2020}. The faintness of WDs proves challenging due to the high signal-to-noise ratio (S/N) required by modern adaptive optics (AO) instruments. This is exacerbated on ground-based telescopes which require extreme AO to counteract the effects of atmospheric seeing. This constrains potential companion searches to relatively bright targets ($m^R\lesssim$ 13). Implicitly, this also limits searches to nearby WDs ($d\sim$\qty{25}{\parsec}, \citealp{holberg25ParsecLocal2016}), which is beneficial for imaging because the limited inner working angle of the instrumentation will project to closer separations around closer stars.

@@

One of the peculiarities of the Sirius system is its large eccentricity of 0.59 \citep{bondSiriusSystemIts2017}. If we assume the orbital expansion due to Sirius B's evolution was adiabatic, we can calculate the initial semi-major axis of the binary
\begin{equation}
    a_i = a_f \frac{M_{B,f} + M_{A,f}}{M_{B,i} + M_{A,i}}
\end{equation}
where $a$ is the system semi-major axis, $M_A$ and $M_B$ are the respective stellar masses, and subscripts $i$ and $f$ correspond to the initial (MS) versus final (post-MS) states \citep{jeansCosmogonicProblemsAssociated1924}. The current semi-major axis of the binary is \qty{20}{\au}, and assuming negligible mass transfer between the two stars, the initial semi-major axis would be \qty{8.6\pm1.3}{\au}. If the orbit expansion was indeed adiabatic, the eccentricity would be the same before and after evolution. In this case, the periastron of Sirius A and B would be \qty{3.52\pm0.52}{\au}. \citet{verasPostmainsequencePlanetarySystem2016} tabulated the maximum stellar radius of intermediate mass stars during their giant evolution, from which we interpolate a maximum radius for Sirius B of \qty{5.104\pm0.075}{\au}. This means Sirius A certainly interacted with Sirius B and may have had a common envelope stage. Mass transfer and tidal circularization would be expected, however the present-day eccentricity provide contrary evidence.

@@

\section{Observations}\label{sec:obs}
\defcitealias{bonnet-bidaudADONISHighContrast2008a}{BB08}
\defcitealias{bondSiriusSystemIts2017}{B17}
\defcitealias{collaborationGaiaEarlyData2021a}{G21a}
\defcitealias{cummingsWhiteDwarfInitialFinal2018}{C18}
\begin{deluxetable}{cccc}
    \tablecaption{Parameters of the Sirius system adopted in this study.\label{tbl:system}}
    \tablehead{
        \colhead{parameter} & \colhead{value} & \colhead{unit} & \colhead{ref.}
    }
    \startdata
    $t_\mathrm{sys}$ & 225 & \unit{\mega\year} & \citetalias{bondSiriusSystemIts2017,cummingsWhiteDwarfInitialFinal2018} \\
    $\pi$ & \num{374.49\pm0.23} & mas & \citetalias{collaborationGaiaEarlyData2021a} \\
    $d$ & \num{2.6702\pm.0016} & pc & \citetalias{collaborationGaiaEarlyData2021a} \\
    $a$ & \num{20.016\pm0.014} & \unit{\au} & \citetalias{bondSiriusSystemIts2017} \\
    $e$ & \num{0.59142\pm0.00037} &  & \citetalias{bondSiriusSystemIts2017} \\
    \cutinhead{Sirius A}
    $M_\star$ & \num{2.063\pm0.023} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017} \\
    \cutinhead{Sirius B}
    $M_\star$ & \num{1.018\pm0.011} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017} \\
    $M_\mathrm{MS}$ & \num{5.1\pm1.1} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017,cummingsWhiteDwarfInitialFinal2018} \\
    $t_\mathrm{WD}$ & 126 & \unit{\mega\year} & \citetalias{bondSiriusSystemIts2017}\\
    $m^{\mathrm{L}^\prime}$ & \num{9.1\pm0.2} & & \citetalias{bonnet-bidaudADONISHighContrast2008a} \\
    $M^{\mathrm{L}^\prime}$ & \num{11.97\pm0.2} & & \citetalias{bonnet-bidaudADONISHighContrast2008a,collaborationGaiaEarlyData2021a} \\
    \enddata
\end{deluxetable}
